### 2022-11-09 paper-thesis/chapters/03-variable-capture.tex
Work on the haskell-coc implementation

The implementation of the base language shown in chapter \ref{chap:baselang} has one big problem, that is name collisions. This section will explore several ways of solving these collisions. An example of such a collision is the following: What is the type of this expression (a polymorphic identity function)?
\begin{lstlisting}
\T : Type. \T : T. T
\end{lstlisting}

@@

\begin{lstlisting}
typeOfExpr(_, FnConstruct("T", Type(), FnConstruct("T", Var("T"), Var("T"))))
\end{lstlisting}

\noindent
This creates a new node in the scope graph, and then type checks the body with this scope.

\begin{lstlisting}
typeOfExpr(s1, FnConstruct("T", Var("T"), Var("T")))

(T : Type)
   [s1]
\end{lstlisting}
TODO Latex library for scope graphs

\noindent
The same thing happens, the body of the \verb|FnConstruct| is typechecked with a new scope. Note that the T in the type of the second T is ambiguous.

\begin{lstlisting}
typeOfExpr(s2, Var("T"))

(T : Type)      (T : Var("T"))
	[s1]<-----------[s2]
\end{lstlisting}

Finally, we need to find the type of \verb|T|. This finds the lexically closest definition of \verb|T| (the one in s2), which is correct. But the type of \verb|T| is \verb|T|, which does NOT refer to the lexically closest \verb|T|, but instead to the \verb|T| in s1. This situation, in which a type can contain a reference to a variable that is shadowed, is the problem. We need to find a way to make sure that shadowing like this can never happen.

>>>>

The implementation of the base language shown in chapter \ref{chap:baselang} has one big problem, that is variable capture. This section will explore several ways of solving this. An example is the following: What is the type of this expression (a polymorphic identity function)?
\begin{lstlisting}
\T : Type. \T : T. T
\end{lstlisting}

@@

\begin{lstlisting}
typeOfExpr(_, FnConstruct("T", Type(), 
    FnConstruct("T", Var("T"), Var("T")))
)
\end{lstlisting}

\noindent
This first handles the outer \verb|FnConstruct|, it creates a new node in the scope graph, and then type checks the body with this scope.

\begin{lstlisting}
typeOfExpr(s1, FnConstruct("T", Var("T"), Var("T")))
\end{lstlisting}
\begin{tikzpicture}[scopegraph, node distance = 2.5em and 3.0em]
\node[scope] (s1) {$s1$};
\node[type, above of = s1] (s1l) {$\texttt{T : Type()}$};
\draw (s1) edge[type] (s1l);
\end{tikzpicture}

\noindent
The same thing happens, the body of the \verb|FnConstruct| is typechecked with a new scope. Note that \verb|Var("T")| in the type of the second \verb|T| is ambiguous, does it refer to the first or second node?

\begin{lstlisting}
typeOfExpr(s2, Var("T"))
\end{lstlisting}
\begin{tikzpicture}[scopegraph, node distance = 2.5em and 2.5em]
	\node[scope] (s1) {$s1$};
	\node[type, above of = s1] (s1l) {$\texttt{T : Type()}$};
	\draw (s1) edge[type] (s1l);

	\node[scope, right of = s1, xshift=5em] (s2) {$s2$};
	\node[type, above of = s2] (s2l) {$\texttt{T : Var("T")}$};
	\draw (s2) edge[type] (s2l);

	\draw (s2) edge[] (s1);
\end{tikzpicture}

Finally, we need to find the type of \verb|Var("T")|. This finds the lexically closest definition of \verb|T| (the one in s2), which is correct. But the type of \verb|T| is \verb|T|, which does NOT refer to the lexically closest \verb|T|, but instead to the \verb|T| in s1. This situation, in which a type can contain a reference to a variable that is shadowed, is the problem. We need to find a way to make sure that shadowing like this can never happen.
